---
title: "Dual Effect of Competition and Predation in the Chilean Rocky Intertidal"
author: "Mary McCabe"
date: "October 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Set your working directory in order to pull in files.
##Add the libraries and packages you plan on using,
```{r}
# Set Working Directory - Connect to Files
setwd("C:/Users/mkmccabe/ECIM Chile/Competition Predation Exp")

suppressPackageStartupMessages({
  library(ggplot2)
  library(tidyr)
  library(dplyr)
  library(DT)
  library(gridExtra)
  library(sciplot)
})
```
##Read in the data files you'll be working with and then combine them, making sure all column line up.
```{r}
# If you have a .csv file (check the original file to know if the separator is ";" or ",").

Temblador=read.csv("Temblador.csv", header=T, sep=",")

Guanaqueros=read.csv("Guanaqueros.csv", header=T, sep=",")

Losmolles=read.csv("Los Molles.csv", header=T, sep=",")

Lascruces=read.csv("Las Cruces.csv", header=T, sep=",")

master <- rbind(Temblador,Guanaqueros,Losmolles,Lascruces)

```

##Add additional columns of entities you wish to explore. In this case I want to look at total coverage of the mussel and barnacle species.
##Change the attribute type of the data column to be in date format.
```{r}
master$PeruProp=(master$Peru + master$Peru.1)/(master$Total.Coverage)*100
master$JehliusProp=(master$Jehlius + master$Jehlius.1)/(master$Total.Coverage)*100
master$NotoProp=(master$Noto + master$Noto.1)/(master$Total.Coverage)*100
master$bothTreat=(paste(master$Depredacion, master$Tratamiento, sep = ""))


master$Fecha=as.Date(master$Fecha, format = "%m/%d/%Y")

datatable(master)  

```

##Clean up data table, include columns you only want to analyze.
%>% Pipe Operator - you think 
Computer shortcut is Ctrl + Shft + m
`select` the columns we need for analysis  
`gather` to move from wide to long - key is associated with column names, value is what is contained in those columns 
`spread` to move from long to wide  
`rename` to rename a column
`mutate` to create a new variables (new columns)
`filter` to filter our data table, maybe by a specific treatment 
```{r}
master_clean <- master %>% 
  rename(Peru_count = n.Peru.Total) %>% 
  select(Sitio, Fecha, Plot, Depredacion, Tratamiento, bothTreat, PeruProp, JehliusProp, NotoProp, Peru_count)
  
datatable(master_clean)  

```

##Number of Mussel Indivuals for each treatment and site and the end of the experiment

```{r}
ind_Peru <- master_clean %>% 
   group_by(Sitio, Fecha, Depredacion, Tratamiento, bothTreat) %>% 
  summarize(mean_count = mean(Peru_count, na.rm = TRUE), se_count = se(Peru_count, na.rm = TRUE))

mussels_only <- ind_Peru %>% 
  filter(bothTreat == "NoM" | bothTreat == "NoMB" | bothTreat == "SiM" | bothTreat == "SiMB")
  

    
head(mussels_only)

```
## Graphing Number of Individual Mussels for each site over time and treatment

```{r}

## Creating my own label with italics as part of the axis label.
mylabel <- c(expression(paste(italic("Perumytilus purpuratus")," Count (# of Individuals)")))


ggplot(mussels_only, aes(Fecha, mean_count, group=bothTreat, shape=bothTreat, colour=bothTreat)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = mean_count - se_count, ymax= mean_count + se_count), width=0.2)+
  theme_classic() +
  labs (x = "Date", y = mylabel) +
  ggtitle ("Spatial and Temporal Variability of Mussel Counts across Treatments")+
  scale_colour_discrete(name="Experimental\nCondition",
                         breaks=c("NoM", "NoMB", "SiM", "SiMB"),
                         labels=c("-P Mono", "-P Mixed", "+P Mono", "+P Mixed")) +
 scale_shape_discrete(name="Experimental\nCondition",
                         breaks=c("NoM", "NoMB", "SiM", "SiMB"),
                         labels=c("-P Mono", "-P Mixed", "+P Mono", "+P Mixed")) +
  scale_x_date(date_labels = "%Y-%m")+
  facet_wrap(~ Sitio, ncol = 2)

  ggsave("musselcounts.pdf")     
  
```




##Mussel Coverage over time

```{r}
cov_Peru <- master_clean %>% 
   group_by(Sitio, Fecha, Depredacion, Tratamiento, bothTreat) %>% 
  summarize(mean_perucov = mean(PeruProp, na.rm = TRUE), se_perucov = se(PeruProp, na.rm = TRUE))

mussels_cov <- cov_Peru %>% 
  filter(bothTreat == "NoM" | bothTreat == "NoMB" | bothTreat == "SiM" | bothTreat == "SiMB")
  

    
head(mussels_cov)

```
## Graphing Coverage of Mussels for each site over time and treatment

```{r}

## Creating my own label with italics as part of the axis label.
mylabel <- c(expression(paste("Proportion of", italic(" P. purpuratus")," (% Coverage)")))


ggplot(mussels_cov, aes(Fecha, mean_perucov, group=bothTreat, shape=bothTreat, colour=bothTreat)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = mean_perucov - se_perucov, ymax= mean_perucov + se_perucov), width=0.2)+
  theme_classic() +
  labs (x = "Date", y = mylabel) +
  ggtitle ("Spatial and Temporal Variability of Mussel Coverage across Treatments")+
  scale_colour_discrete(name="Experimental\nCondition",
                         breaks=c("NoM", "NoMB", "SiM", "SiMB"),
                         labels=c("-P Mono", "-P Mixed", "+P Mono", "+P Mixed")) +
 scale_shape_discrete(name="Experimental\nCondition",
                         breaks=c("NoM", "NoMB", "SiM", "SiMB"),
                         labels=c("-P Mono", "-P Mixed", "+P Mono", "+P Mixed")) +
  scale_x_date(date_labels = "%Y-%m")+
  facet_wrap(~ Sitio, ncol = 2)

  ggsave("musselcoverage.pdf", width = 20, height = 13, units = "cm")     
  
```


##Barnacle (Jehlius) Coverage over time

```{r}
cov_Jehlius <- master_clean %>% 
   group_by(Sitio, Fecha, Depredacion, Tratamiento, bothTreat) %>% 
  summarize(mean_jehcov = mean(JehliusProp, na.rm = TRUE), se_jehcov = se(JehliusProp, na.rm = TRUE))

jehlius_cov <- cov_Jehlius %>% 
  filter(bothTreat == "NoB" | bothTreat == "NoMB" | bothTreat == "SiB" | bothTreat == "SiMB")
  

    
head(jehlius_cov)

```

## Graphing Coverage of (Jehlius) Barnacles for each site over time and treatment

```{r}

## Creating my own label with italics as part of the axis label.
mylabel <- c(expression(paste("Proportion of", italic(" J. cirratus")," (% Coverage)")))


ggplot(jehlius_cov, aes(Fecha, mean_jehcov, group=bothTreat, shape=bothTreat, colour=bothTreat)) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymin = mean_jehcov - se_jehcov, ymax= mean_jehcov + se_jehcov), width=0.2)+
  theme_classic() +
  labs (x = "Date", y = mylabel) +
  ggtitle ("Spatial and Temporal Variability of Jehlius cirratus Coverage across Treatments")+
  scale_colour_discrete(name="Experimental\nCondition",
                         breaks=c("NoB", "NoMB", "SiB", "SiMB"),
                         labels=c("-P Mono", "-P Mixed", "+P Mono", "+P Mixed")) +
 scale_shape_discrete(name="Experimental\nCondition",
                         breaks=c("NoB", "NoMB", "SiB", "SiMB"),
                         labels=c("-P Mono", "-P Mixed", "+P Mono", "+P Mixed")) +
  scale_x_date(date_labels = "%Y-%m")+
  facet_wrap(~ Sitio, ncol = 2)

  ggsave("barnaclecoverage.pdf", width = 20, height = 13, units = "cm")     
  
```






